---
title: "Machine Learning Project"
output: html_document
---

## Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).


## Analysis
### Approach
* Load the data set and briefly learn the characteristics of the data
* Use cross-validation method to built a valid model; 70% of the original data is used for model building (training data) while the rest of 30% of the data is used for testing (testing data)
* Since the number of variables in the training data is too large, clean the data by 1) excluding variables which apparently cannot be explanatory variables, and 2) reducing variables with little information.
* Apply PCA to reduce the number of variables
* Apply random forest method to build a model
* Check the model with the testing data set
* Apply the model to estimate classes of 20 observations


### Loading data
```{r}
data <- read.csv("pml-training.csv")
colnames(data)
summary(data)
```

### Cross validation
Use 70% of training set data to built a model, and use the rest to test the model
```{r}
library(caret)
```

```{r}
set.seed(1111)
train <- createDataPartition(y=data$classe,p=.70,list=F)
training <- data[train,]
testing <- data[-train,]
```

### Cleaning the training data
Exclude identifier, timestamp, and window data
```{r}
Cl <- grep("name|timestamp|window|X", colnames(training), value=F) 
trainingCl <- training[,-Cl]
```

Select variables with high (over 95%) missing data, exclude them from the analysis
```{r}
trainingCl[trainingCl==""] <- NA
NArate <- apply(trainingCl, 2, function(x) sum(is.na(x)))/nrow(trainingCl)
trainingCl <- trainingCl[!(NArate>0.95)]
summary(trainingCl)
```

### PCA
Since the number of variables are still over 50, PCA is applied
```{r}
preProc <- preProcess(trainingCl[,1:52],method="pca",thresh=.8) #12 components are required
preProc <- preProcess(trainingCl[,1:52],method="pca",thresh=.9) #18 components are required
preProc <- preProcess(trainingCl[,1:52],method="pca",thresh=.95) #25 components are required

preProc <- preProcess(trainingCl[,1:52],method="pca",pcaComp=25) 
preProc$rotation
trainingPC <- predict(preProc,trainingCl[,1:52])
```

### Random forest
Apply ramdom forest method (non-bionominal outcome & large sample size)
```{r}
library(randomForest)
```

```{r}
modFitRF <- randomForest(trainingCl$classe ~ .,   data=trainingPC, do.trace=F)
print(modFitRF)
```

Importance of each predictor.
```{r}
importance(modFitRF)
```

### Check with test set
```{r}
testingCl <- testing[,-Cl]
testingCl[testingCl==""] <- NA
NArate <- apply(testingCl, 2, function(x) sum(is.na(x)))/nrow(testingCl)
testingCl <- testingCl[!(NArate>0.95)]
testingPC <- predict(preProc,testingCl[,1:52])
confusionMatrix(testingCl$classe,predict(modFitRF,testingPC))
```

### Predict classes of 20 test data
```{r}
testdata <- read.csv("pml-testing.csv")
testdataCl <- testdata[,-Cl]
testdataCl[testdataCl==""] <- NA
NArate <- apply(testdataCl, 2, function(x) sum(is.na(x)))/nrow(testdataCl)
testdataCl <- testdataCl[!(NArate>0.95)]
testdataPC <- predict(preProc,testdataCl[,1:52])
testdataCl$classe <- predict(modFitRF,testdataPC)
```


## Conclusion
In this analyses, 19622 observations from weight lifting exercise were used to analyze and predict correct body movement from others during the exercise. 70% of the total observations (13737 observations) were used to build a model by random forest method, and the rest of 30% of the observations (5885 observations) were used for model validation (cross-validation). The model statistics showed that the built model had the overall accuracy of 97% for the testing set, which is not overlapping with observations used to built the model. The sensitivity was in between 92%-99% and the specificity was over 99% for all classes (class A-E, total 5 classes. class A is the data from correct exercise while the other classes were data from exercises done in a wrong way). Overall, the model is well developed to predict the exercise classes during weight lifting. As for the limitation in this study, the observation data used in the analyses was collected from 6 young health participants in an experiment using Microsoft Kinect. Therefore, under those condition, the model is expected to perform over 95% accuracy; however, with different conditions, such as experiments with elderly people and/or using different device, the model might not perform well as shown in the analysis.